---
title: "REDRESS - District sampling"
author: "Lucas Sempe & Yan Ding"
date: "31/01/2022"
output:
  bookdown::gitbook:
    highlight: espresso
    number_sections: false
header-includes:
   - \usepackage{booktabs}
   - \usepackage{multirow}
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,message = F)



```

```{r, warning=FALSE, message=F, echo=FALSE}

source("draw.R")


texPreview::tex_opts$set( 
  returnType = 'html',
  opts.html   = list(width = '75%', height = '75%')
)

```

## Introduction

The purpose of this document is set the guidelines and suggest a
sampling strategy for the implementation and evaluation of the REDRESS
project.

The study of the intervention and the evaluation will collect primary
data exclusively in the three treatment counties: **Lofa** for ***Study
A***, and **GrandGedeh** and **Margibi** for ***Study B***.

There is an agreement between partners to select two districts of each
county, which will allow to capture geographical and socio-economic
diversity.

## Criteria

We suggest the combination of following criteria to select the
districts:

-   Urban/rural density in order to reflect different health and social
    dynamics;

-   Larger number of health clinics in order to have a larger pool of
    potential health facilities to survey;

-   Districts with larger number of SSSD cases registered in 2019 and
    2020, understanding that prevalence is relatively low in population
    and there is a need to capture a significant number of observations
    to gain as much as statistical power as possible.

## Prevalence

Considering there is are relatively low number of SSSDs, we review
evidence of prevalence of the diseases in recent years. @simpson2019 did
a systematic review of **Buruli ulcer** prevalence worldwide. Prevalence
estimates within study areas ranged from 3.2 (95% CI 3.1-3.3) cases per
10000 population in Côte d'Ivoire to 26.9 [95% CI 23.5-30.7] cases per
10,000 population in Benin. There was evidence of Buruli ulcer in 32
countries and consensus on presence in 12.

@timothy2021 conducted a population-based cross-sectional integrated
survey for skin NTDs (Buruli ulcer, leprosy, lymphatic
filariasis--associated morbidity, and yaws) during June--October 2018 in
Maryland County, Liberia (census population 165,456). The find confirmed
endemicity and autochthonous transmission of **yaws** in Liberia after a
population-based, community-led burden estimation (56,825 participants),
24 cases; 2.6 [95% CI 1.4-3.9] cases/10,000 population). Aside from 1
case in a 32-year-old person, all confirmed cases were in persons <18
years of age, most of whom were male.

@cromwell2020 estimate a posterior probability that the prevalence of
**lymphatic filariasis** by immunochromatographic test as of 2018 was
unlikely to be below 1% in Ghana, Liberia, northern Democratic Republic
of the Congo, and Côte d'Ivoire. [@who2021] reports 109 new cases of **leprosis** (registered prevalence) in Liberia in 2019. Finally,[@vinkelesmelchers2021] estimate a prevalence lower than <0.05% for **onchocerciasis** in Liberia.

Data from the Ministry of Health (MoH) of Liberia also provides additional insights. Table \@ref(tab:raw) shows the number of cases from the selected diseases from 2017 to 2019. We observe that years 2017 and 2018 show a larger number of reported cases, where leprosis, buruli ulcer and lymphatic filariasis show the higher number of cases at country level. It is worth noticing that at county level, the highest single value found is 42 cases of buruli ulcer in Lofa in 2017.

|      	|      	| Margibi 	| G. Gedeh 	| Lofa 	|  National Annual Total  	|
|------	|------	|---------	|----------	|------	|:-----------------------:	|
| 2017 	| LEPR 	| 0       	| 12       	| 0    	| 164                     	|
|      	| BU   	| 20      	| 9        	| 42   	| 202                     	|
|      	| HY   	| 3       	| 0        	| 0    	| 122                     	|
|      	| LY   	| 4       	| 1        	| 9    	| 223                     	|
|      	| YAWS 	| 0       	| 0        	| 0    	| 0                       	|
| 2018 	| LEPR 	| 1       	| 2        	| 3    	| 189                     	|
|      	| BU   	| 29      	| 4        	| 21   	| 324                     	|
|      	| HY   	| 0       	| 0        	| 6    	| 25                      	|
|      	| LY   	| 0       	| 0        	| 11   	| 64                      	|
|      	| YAWS 	| 0       	| 0        	| 0    	| 19                      	|
| 2019 	| LEPR 	| 2       	| 4        	| 1    	| 62                      	|
|      	| BU   	| 3       	| 2        	| 20   	| 129                     	|
|      	| HY   	| 0       	| 0        	| 0    	| 0                       	|
|      	| LY   	| 0       	| 0        	| 1    	| 18                      	|
|      	| YAWS 	| 0       	| 0        	| 5    	| 28                      	|

: (#tab:raw) Number of cases in selected counties and national level by SSSDs - 2017 to 2019

Table \@ref(tab:incidence) shows the computation of the incidence of new cases by 10,000 inhabitants. This suggests, if compared to the previously aforementioned prevalence/incidence data a potential combination of under-reporting and low prevelance.

|      | Margibi | G. Gedeh | Lofa | National Annual Total |
|------|---------|----------|------|:---------------------:|
| LEPR | 0.00    | 0.02     | 0.00 |         0.33          |
| BU   | 0.04    | 0.02     | 0.08 |         0.40          |
| HY   | 0.01    | 0.00     | 0.00 |         0.24          |
| LY   | 0.01    | 0.00     | 0.02 |         0.45          |
| YAWS | 0.00    | 0.00     | 0.00 |         0.00          |
| LEPR | 0.00    | 0.00     | 0.01 |         0.38          |
| BU   | 0.06    | 0.01     | 0.04 |         0.65          |
| HY   | 0.00    | 0.00     | 0.01 |         0.05          |
| LY   | 0.00    | 0.00     | 0.02 |         0.13          |
| YAWS | 0.00    | 0.00     | 0.00 |         0.04          |
| LEPR | 0.00    | 0.01     | 0.00 |         0.12          |
| BU   | 0.01    | 0.00     | 0.04 |         0.26          |
| HY   | 0.00    | 0.00     | 0.00 |         0.00          |
| LY   | 0.00    | 0.00     | 0.00 |         0.04          |
| YAWS | 0.00    | 0.00     | 0.01 |         0.06          |

: (#tab:incidence) Incidence by 10,000 inhabitants in selected counties and national level by SSSDs - 2017 to 2019

## Sample proposal

In the following tables, we present summary statistics for the three
intervention counties. In each case, we argue in favour of selecting
certain districts based on the aforementioned criteria.

Table \@ref(tab:lofa) shows 7 districts in Lofa County, where we
identify two potential districts: Foya and Voinjama (where the County
capital is located) are the ones with larger number of health facilities
(3rd and 1st), among the largest share of population covered (1st and
3rd, respectively) and the largest number of SSSD cases identified in
2019 and 2020 (1st and 2nd, respectively).

An alternative could be the Kolahum District. However, we found news
that suggest the district has been partitioned into smaller
portions[^1], although we could no find official information the
boundaries.

[^1]: <https://allafrica.com/stories/201509281614.html>

```{r lofa, fig.cap="Summary statistics for districts in Lofa county"}

lof<-Sampling_frame %>% filter (County=="Lofa") %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))
loft<-flextable::flextable(lof)

loft <- set_caption(loft, "District: Lofa")

loft

```

Table \@ref(tab:margibi) shows 4 districts in Margibi County. We find
three potential districts. First, the Kakata District, where the capital is
located and it is the most populated and with the largest number of
health facilities - although there where only 1 SSSD in 2019. Second, the
Mambah-Kaba District, which is second in number of health facilities and
first in number of cases. Finally, the Firestone District appears as a candidate due to the larger relative population and the highster number of registered cases in 2020.

```{r margibi, fig.cap="Summary statistics for districts in Margibi county"}

marg<-Sampling_frame %>% filter (County == "Margibi" ) %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))
marg.t<-flextable::flextable(marg)

marg.t <- set_caption(marg.t, "District: Margibi")

marg.t

```

Table \@ref(tab:grand) presents districts from the Grand Gedeh county, where the Tchien district appears as the best candidate in terms of higher population, health facilities and SSSD cases identified in 2020. The other districts do not show very relevant differences in terms of population and SSDD cases. Based on that, we suggest choosing the Cavalla District as it has the largest pool of health facilities (6) among the remaining districts.

```{r grand, fig.cap="Summary statistics for districts in Grand Gedeh county"}

gg<-Sampling_frame %>% filter (County == "Grand Gedeh") %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))

gg.t<-flextable::flextable(gg)

gg.t <- set_caption(gg.t, "District: Grand Gedeh")

gg.t


```

## Map

The following map presents 89 Health Facilities mapped by USAID in
2014/2015, that where matched to the database REDRESS received. There
are 40 additional Health Facilities in the REDRESS database (n=139) that
could not be geolocated.

```{r map, fig.cap="Map witn intervention Districs and health facilities"}

m1

```

## References
