---
title: "REDRESS - District sampling"
author: "Lucas Sempe & Yan Ding"
date: "31/01/2022"
output: html_document
header-includes:
   - \usepackage{booktabs}
   - \usepackage{multirow}
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, warning=FALSE, message=F, echo=FALSE}

source("draw.R")


```

## Introduction

The purpose of this document is set the guidelines and suggest a sampling strategy for the implementation and evaluation of the REDRESS project.

The study of the intervention and the evaluation will collect primary data exclusively in the three treatment counties: **Lofa** for ***Study A***, and **GrandGedeh** and **Margibi** for ***Study B***.

There is an agreement between partners to select two districts of each county, which will allow to capture geographical and socio-economic diversity.

## Criteria

We suggest the combination of following criteria to select the districts:

-   Urban/rural density in order to reflect different health and social dynamics;

-   Larger number of health clinics in order to have a larger pool of potential health facilities to survey;

-   Districts with larger number of SSSD cases registered in 2019 and 2020, understanding that prevalence is relatively low in population and there is a need to capture a significant number of observations to gain as much as statistical power as possible.

## Prevalence

Considering there is are relatively low number of SSSDs, we review evidence of prevalence of the diseases in recent years. @simpson2019 did a systematic review of **Buruli ulcer** prevalence worldwide. Prevalence estimates within study areas ranged from 3·2 (95% CI 3·1--3·3) cases per 10 000 population in Côte d'Ivoire to 26·9 (23·5--30·7) cases per 10 000 population in Benin. There was evidence of Buruli ulcer in 32 countries and consensus on presence in 12.

@timothy2021 conducted a population-based cross-sectional integrated survey for skin NTDs (Buruli ulcer, leprosy, lymphatic filariasis--associated morbidity, and yaws) during June--October 2018 in Maryland County, Liberia (census population 165,456). The find confirmed endemicity and autochthonous transmission of **yaws** in Liberia after a population-based, community-led burden estimation (56,825 participants), 24 cases; 2.6 \[95% CI 1.4--3.9\] cases/10,000 population). Aside from 1 case in a 32-year-old person, all confirmed cases were in persons [\<]{.ul}18 years of age, most of whom were male.

@cromwell2020 estimate a posterior probability that the prevalence of **lymphatic filariasis** by immunochromatographic test as of 2018 was unlikely to be below 1% in Ghana, Liberia, northern Democratic Republic of the Congo, and Côte d'Ivoire.

[@ohanlon2016]

    \begin{table}[]
    \begin{tabular}{@{}clllll@{}}
    \toprule
    \multicolumn{1}{l}{\textbf{}}  & \textbf{}     & \textbf{Margibi} & \textbf{G. Gedeh} & \textbf{Lofa} & \multicolumn{1}{c}{\textbf{National Annual Total}} \\ \midrule
    \multirow{5}{*}{\textbf{2017}} & \textbf{LEPR} & 0                & 12                & 0             & 164                                                \\
                                   & \textbf{BU}   & 20               & 9                 & 42            & 202                                                \\
                                   & \textbf{HY}   & 3                & 0                 & 0             & 122                                                \\
                                   & \textbf{LY}   & 4                & 1                 & 9             & 223                                                \\
                                   & \textbf{YAWS} & 0                & 0                 & 0             & 0                                                  \\
    \multirow{5}{*}{\textbf{2018}} & \textbf{LEPR} & 1                & 2                 & 3             & 189                                                \\
                                   & \textbf{BU}   & 29               & 4                 & 21            & 324                                                \\
                                   & \textbf{HY}   & 0                & 0                 & 6             & 25                                                 \\
                                   & \textbf{LY}   & 0                & 0                 & 11            & 64                                                 \\
                                   & \textbf{YAWS} & 0                & 0                 & 0             & 19                                                 \\
    \multirow{5}{*}{\textbf{2019}} & \textbf{LEPR} & 2                & 4                 & 1             & 62                                                 \\
                                   & \textbf{BU}   & 3                & 2                 & 20            & 129                                                \\
                                   & \textbf{HY}   & 0                & 0                 & 0             & 0                                                  \\
                                   & \textbf{LY}   & 0                & 0                 & 1             & 18                                                 \\
                                   & \textbf{YAWS} & 0                & 0                 & 5             & 28                                                 \\ \cmidrule(l){2-6} 
    \end{tabular}
    \end{table}

## Sample proposal

In the following tables, we present summary statistics for the three intervention counties. In each case, we argue in favour of selecting certain districts based on the aforementioned criteria.

Table \@ref(tab:lofa) shows 7 districts in Lofa County, where we identify two potential districts: Foya and Voinjama (where the County capital is located) are the ones with larger number of health facilities (3rd and 1st), among the largest share of population covered (1st and 3rd, respectively) and the largest number of SSSD cases identified in 2019 and 2020 (1st and 2nd, respectively).

An alternative could be the Kolahum District. However, we found news that suggest the district has been partitioned into smaller portions[^1], although we could no find official information the boundaries.

[^1]: <https://allafrica.com/stories/201509281614.html>

```{r lofa, fig.cap="Summary statistics for districts in Lofa county"}

lof<-Sampling_frame %>% filter (County=="Lofa") %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))
loft<-flextable::flextable(lof)

loft <- set_caption(loft, "District: Lofa")

loft

```

Table \@ref(tab:margibi) shows 4 districts in Margibi County. We find three potential districts. First, Kakata District, where the capital is located and it is the most populated and with the largest number of health facilities - although there where only 1 SSSD in 2019. Second, Mambah-Kaba District, which is second in number of health facilities and first in number of cases

```{r margibi, fig.cap="Summary statistics for districts in Margibi county"}

marg<-Sampling_frame %>% filter (County == "Margibi" ) %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))
marg.t<-flextable::flextable(marg)

marg.t <- set_caption(marg.t, "District: Margibi")

marg.t

```

```{r grand, fig.cap="Summary statistics for districts in Grand Gedeh county"}

gg<-Sampling_frame %>% filter (County == "Grand Gedeh") %>%
  group_by(County,District) %>% summarise (`# Health Facilities`=n(),
                                           `Population covered`=sum(`Population covered`,na.rm = T),
                                           `# SSSD cases 2020` = sum(`SSSD cases 2020 (total)`,na.rm=T
                                           ),
                                           `# SSSD cases 2019` = sum(`SSSD cases 2019 (total)`,na.rm=T
                                           ))

gg.t<-flextable::flextable(gg)

gg.t <- set_caption(gg.t, "District: Grand Gedeh")

gg.t


```

## Map

The following map presents 89 Health Facilities mapped by USAID in 2014/2015, that where matched to the database REDRESS received. There are 40 additional Health Facilities in the REDRESS database (n=139) that could not be geolocated.

```{r map, fig.cap="Summary statistics for districts in Lofa county"}

m1

```
